---
title: "BCA_20240617"
author: "Christoph Schweingruber"
output: html_document
date: "2024-06-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(openxlsx)
library(ggprism)
library(renv)
```

## BCA assay

```{r load_data}
source("R/read_clariostar.R")
data <- read_clariostar(file = "data/BCA_20240617_1.CSV")
meta <- read_excel(path = "meta/Samples_BCA_20240617.xlsx") %>%
  select(.data = ., -Column, - Row)

data <- left_join(
  x = meta,
  y = data,
  by = c("Well" = "Well")
) %>%
  janitor::clean_names(.)

#data
```

The manual for the Pierce BCA Protein Assay suggest that a linear regression 
might be worse than non-linear regression for the calibration curve. It suggests
to use a 'four-parameter quadratic' equation, which is commonly called a
logistic regression. Let's use it here:

```{r standard_curve}
source("R/calibrate_assay.R")

standards <- filter(data, task == "Standard")
calibration_curve <- fit_logistic_calibration_curve(standards, level = 0.95)

plot1 <- ggplot(data = standards, aes(x = quantity, y = a562)) +
  geom_point() +
  geom_ribbon(
    data = calibration_curve$fit,
    mapping = aes(x = quantity, ymin = lwr, ymax = upr),
    fill = "dodgerblue",
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  geom_line(
    data = calibration_curve$fit,
    mapping = aes(x = quantity, y = fit),
    linetype = "dashed",
    inherit.aes = FALSE
  ) +
  annotate(
    geom = "text",
    label = 'paste(psi, {R^2}["adj"], " = ")',
    x = 0,
    y = 2,
    parse = TRUE,
    hjust = "left"
  ) +
  geom_text(
    data = calibration_curve$r_squared,
    mapping = aes(x = 0.28, y = 2, label = round(adj_R_squared, 4))
  ) +
  labs(
    x = "Albumin, mg/mL",
    y = "Absorbance A562",
    title = "Calibration curve",
    subtitle = "Fit + CI(95%)"
  ) +
  theme_prism()

plot1
```

As we can see, the fit is okay.

```{r quantify_samples}
source("R/calibrate_assay.R")
data <- quantify_unknowns(data, calibration_curve)

unknowns <- data %>%
  filter(., task %in% c("Unknown", "Negative"))

plot2 <- ggplot(data = standards, aes(x = a562, y = quantity)) +
  geom_ribbon(
    data = calibration_curve$fit,
    mapping = aes(y = quantity, xmin = lwr, xmax = upr),
    fill = "dodgerblue",
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  geom_line(
    data = calibration_curve$fit,
    mapping = aes(x = fit, y = quantity),
    linetype = "dashed",
    inherit.aes = FALSE
  ) +
  geom_point() +
  geom_point(
    data = unknowns,
    mapping = aes(color = clone)
  ) +
  geom_rug(
    data = unknowns,
    mapping = aes(color = clone)
  ) +
  scale_color_manual(values = hcl.colors(6, palette = "Sunset")) +
  annotate(
    geom = "text",
    label = 'paste(psi, {R^2}["adj"], " = ")',
    x = -0.2,
    y = 2,
    parse = TRUE,
    hjust = "left"
  ) +
  geom_text(
    data = calibration_curve$r_squared,
    mapping = aes(x = 0.3, y = 2, label = round(adj_R_squared, 4))
  ) +
  labs(
    x = "Absorbance A562",
    y = "Protein, mg/mL",
    title = "BCA protein assay",
    subtitle = "calibration curve + CI(95%)"
  ) +
  theme_prism()

plot2
```

Our samples are in the lower range but reasonable.


```{r write_data}
wb <- createWorkbook()

addWorksheet(wb, sheetName = "Data")
freezePane(wb, sheet = "Data", firstRow = TRUE, firstCol = TRUE)
writeDataTable(
  wb,
  sheet = "Data",
  x = data,
  colNames = TRUE,
  rowNames = FALSE,
  tableStyle = "None"
)
addWorksheet(wb, sheetName = "Summary")
freezePane(wb, sheet = "Summary", firstRow = TRUE, firstCol = TRUE)
writeDataTable(
  wb,
  sheet = "Summary",
  x = select(
    .data = arrange(.data = unknowns, clone, shield),
    well, name, task, clone, shield, a562, quantity, unit
  ),
  colNames = TRUE,
  rowNames = FALSE,
  tableStyle = "None"
)

## Save workbook
saveWorkbook(wb, "docs/BCA_20240617.xlsx", overwrite = TRUE)
```


## References

* https://bookdown.org/hneth/ds4psy/D-3-apx-colors-basics.html
* https://www.myassays.com/four-parameter-logistic-regression.html


